---
title: "Carcass Recovery"
author: "jim.carretta@noaa.gov"
date: '2022-06-02'
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What fraction of cetacean carcasses are documented?

Estimating the proportion of human-caused mortality and serious injury to marine mammals
that is documented requires three general data sets:

1. The number of stranded and/or floating carcasses detected.
2. Estimated population size for the species of interest.
3. Annual survival estimates, ideally including anthropogenic impacts.

It is then straightforward to incorporate these data into models used to estimate total annual deaths expected:

```{r}
PopSize <- 10000
AnnualSurvival <- 0.98

AnnualDeaths <- PopSize * (1 - AnnualSurvival)
AnnualDeaths

ObservedDeaths <- 8

FractionDetected <- ObservedDeaths/AnnualDeaths
FractionDetected

```



## References

[Carretta, J. V., Danil, K., Chivers, S. J., Weller, D. W., Janiger, D. S., Berman-Kowalewski, M., … Lambourn, D. M. (2016). Recovery rates of bottlenose dolphin (Tursiops truncatus) carcasses estimated from stranding and survival rate data. Marine Mammal Science, 32, 349– 362.](https://doi.org/10.1111/mms.12264)

[Harting, A.L., Michelle M. Barbieri, Jason D. Baker, Tracy A. Mercer, Thea C. Johanos, Stacie J. Robinson, Charles L. Littnan, Katie M. Colegrove, and Dave S. Rotstein. (2020). Population‐level impacts of natural and anthropogenic causes‐of‐death for Hawaiian monk seals in the main Hawaiian Islands, Marine Mammal Science 37(1):235-250](https://onlinelibrary.wiley.com/doi/full/10.1111/mms.12742).

[Pace III, R.M., Williams, R., Kraus, S.D., Knowlton, A.R. and Pettis, H.M., 2021. Cryptic mortality of North Atlantic right whales. Conservation Science and Practice, 3(2), p.e346.](https://conbio.onlinelibrary.wiley.com/doi/full/10.1111/csp2.346)

[Williams, R., Gero, S., Bejder, L., Calambokidis, J., Kraus, S.D., Lusseau, D., Read, A.J. and Robbins, J., 2011. Underestimating the damage: interpreting cetacean carcass recoveries in the context of the Deepwater Horizon/BP incident. Conservation Letters, 4(3), pp.228-233.](https://doi.org/10.1111/j.1755-263X.2011.00168.x)

[Wells, R. S., J. B. Allen, G. Lovewell, J. Gorzelany, R. E. Delynn, D. A. Fauquier and N. B. Barros. 2015. Carcass-recovery rates for resident bottlenose dolphins in Sarasota Bay, Florida. Marine Mammal Science 31:355–368.](https://onlinelibrary.wiley.com/doi/10.1111/mms.12142)



