---
title: "Carcass Recovery"
author: "jim.carretta@noaa.gov"
date: '2022-06-02'
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What fraction of cetacean carcasses are documented?

Estimating the proportion of human-caused mortality and serious injury to marine mammals that is documented requires several data types:

1.    Abundance
2.    Annual Survival Rate (ASR) or 1 - Natural Mortality Rate
3.    Number of Recovered Carcasses / Strandings
4.    Fraction of year spent in study area (e.g. blue whale occurrence in U.S. West Coast waters = 0.583 or 7/12 Yr; Hazen et al. 2016)
5.    Estimated annual anthropogenic removals

It is then straightforward to incorporate these data into models used to estimate total annual deaths expected in a given area.
Using a U.S. West Coast blue whale example:

```{r}
PopSize <- 2000
AnnualSurvival <- 0.98
FrResidencyTime <- 0.583
AnthropogenicRemovals <- 2

AnnualDeaths <- PopSize * (1 - AnnualSurvival) * FrResidencyTime + AnthropogenicRemovals
AnnualDeaths

ObservedDeaths <- 4

FractionDetected <- ObservedDeaths/AnnualDeaths
FractionDetected

```



## References

[Carretta, J. V., Danil, K., Chivers, S. J., Weller, D. W., Janiger, D. S., Berman-Kowalewski, M., … Lambourn, D. M. (2016). Recovery rates of bottlenose dolphin (Tursiops truncatus) carcasses estimated from stranding and survival rate data. Marine Mammal Science, 32, 349– 362.](https://doi.org/10.1111/mms.12264)

[Harting, A.L., Michelle M. Barbieri, Jason D. Baker, Tracy A. Mercer, Thea C. Johanos, Stacie J. Robinson, Charles L. Littnan, Katie M. Colegrove, and Dave S. Rotstein. (2020). Population‐level impacts of natural and anthropogenic causes‐of‐death for Hawaiian monk seals in the main Hawaiian Islands, Marine Mammal Science 37(1):235-250](https://onlinelibrary.wiley.com/doi/full/10.1111/mms.12742).

[Hazen, E.L., D.M. Palacios, K.A. Forney, E.A. Howell, E. Becker, A.L. Hoover, L. Irvine, M. DeAngelis, S.J. Bograd, B.R. Mate, H. Bailey. 2016. WhaleWatch: a dynamic management tool for predicting blue whale density in the California Current. Journal of Applied Ecology](https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/1365-2664.12820)

[Pace III, R.M., Williams, R., Kraus, S.D., Knowlton, A.R. and Pettis, H.M., 2021. Cryptic mortality of North Atlantic right whales. Conservation Science and Practice, 3(2), p.e346.](https://conbio.onlinelibrary.wiley.com/doi/full/10.1111/csp2.346)

[Read, A.J., Drinker, P. and Northridge, S., 2006. Bycatch of marine mammals in US and global fisheries. Conservation biology, 20(1), pp.163-169.](https://conbio.onlinelibrary.wiley.com/doi/10.1111/j.1523-1739.2006.00338.x)

[Williams, R., Gero, S., Bejder, L., Calambokidis, J., Kraus, S.D., Lusseau, D., Read, A.J. and Robbins, J., 2011. Underestimating the damage: interpreting cetacean carcass recoveries in the context of the Deepwater Horizon/BP incident. Conservation Letters, 4(3), pp.228-233.](https://doi.org/10.1111/j.1755-263X.2011.00168.x)

[Wells, R. S., J. B. Allen, G. Lovewell, J. Gorzelany, R. E. Delynn, D. A. Fauquier and N. B. Barros. 2015. Carcass-recovery rates for resident bottlenose dolphins in Sarasota Bay, Florida. Marine Mammal Science 31:355–368.](https://onlinelibrary.wiley.com/doi/10.1111/mms.12142)



